Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-10-04 paper/paper.tex
added another section to discussion

\section{Discussion}
We have determined the full sequences of the infecting virions in single influenza-infected cells, and then examined how viral mutations and gene-expression defects associate with infection outcome in each cell.
Methodologically, the major advance of our work is to measure the \emph{genotypes} of viruses in single cells in addition to quantifying the abundance of viral components (i.e., transcripts, proteins, or progeny virions) as has been done by prior single-cell virology studies~\citep{russell2018extreme, zanini2018single, zanini2018virus, steuerman2018dissection, saikia2018simultaneous, zhu2009growth, schulte2014single, akpinar2016high, heldt2015single, brooke2013most}.
Our method builds on the observations by \citet{saikia2018simultaneous} and \citet{zanini2018virus} that fragmentary viral genetic information can be obtained using standard single-cell transcriptomic techniques.
To make this genetic information complete, we have coupled a standard single-cell transcriptomic technique to long-read PacBio sequencing, a strategy analogous to that recently used by \cite{gupta2018single} to obtain full-length isoforms of some cellular genes in single cells.

Our results demonstrate that the single-cell viral genetic information provided by this methodological advance is extremely valuable for understanding the biology of influenza infection, especially as it relates to innate-immune detection.

>>>>

\section{Discussion}
We have determined the full sequences of the viral genes expressed in single influenza-infected cells, and examined how viral mutations and gene-expression defects associate with infection outcome in each cell.
Methodologically, our major advance is to measure the \emph{genotypes} of viruses in addition to measuring the abundance of viral components (i.e., transcripts, proteins, or progeny virions) as has been done by prior single-cell studies~\citep{russell2018extreme, zanini2018single, zanini2018virus, steuerman2018dissection, saikia2018simultaneous, zhu2009growth, schulte2014single, akpinar2016high, heldt2015single, brooke2013most}.
Our method builds on the observations by \citet{saikia2018simultaneous} and \citet{zanini2018virus} that fragmentary viral genetic information can be obtained using more standard single-cell transcriptomic techniques.
To make this information complete, we have coupled a standard single-cell transcriptomic technique to long-read PacBio sequencing of viral genes, a strategy analogous to that recently used by \cite{gupta2018single} to obtain full-length isoforms of some cellular genes in single cells.

This viral genetic information turns out to be extremely important for understanding infection outcome and innate-immune induction.
Despite the fact that we used a low-passage viral stock generated from plasmids encoding a ``wildtype'' influenza genome, a large majority of cells are infected by virions that have mutations or fail to express a gene.
Although our study is certainly not the first to note that influenza virus has a high mutation rate~\citep{parvin1986measurement, suarez1992heterogeneity, suarez1994estimation, bloom2014experimentally, pauly2017novel} and sometimes fails to express genes~\citep{brooke2013most, heldt2015single, dou2017analysis, russell2018extreme}, it is the first to directly observe the full spectrum of these defects across single cells.
Simple visual inspection of \FIG{genotypes} makes it glaringly obvious how any experimental approach that does not directly sequence the viral genes in single cells is really averaging across a vast spectrum of hidden viral defects.

\jdbcomment{Our results demonstrate that the single-cell viral genetic information provided by this methodological advance is extremely valuable for understanding the biology of influenza infection, especially as it relates to innate-immune detection.}
